plyranges: a grammar for transforming genomics data

### 2018-03-22 paper.Rmd
a more abstract definition of the Ranges

<--- need to formally introduce the Ranges classes now --->

>>>>

The `plyranges` DSL is built on two fundamental data structures defined
by the `IRanges` and `GenomicRanges` packages called _IntegerRanges_
and _GenomicRanges_. Unless otherwise specified, throughout the paper
we refer to either of these data structures as _Ranges_. In its simplest form 
the data model for an _IntegerRanges_ object is defined as a table with
$n$ rows, where each row contains a closed interval of two integers:

$$
\textit{IntegerRange} := \{ \left[\text{start}_i, \text{end}_{i}\right]  : i = 1\dots n \}.
$$
In this form an _IntegerRanges_ object has three core columns: a start, an
end, and the width of the interval. The _GenomicRanges_ data
model has additional semantics providing biological context, strand and 
sequence name (seqname), as well as containing an _IntegerRanges_ object.

$$
\textit{GenomicRange} := \{  \{\text{seqname}_i, \left[\text{start}_i, \text{end}_{i}\right] , \text{strand}_i\}  : i = 1\dots n \}.
$$
A _GenomicRanges_ object can have an additional property that provides
annotation for each sequence name over a genome. This contstrains the the 
start and end points for a given sequence to lie inside the bounds
of the genome space. Additional values can be bound to the bare _IntegerRanges_ or _GenomicRanges_ and form metadata columns.

By definition the _Ranges_ class encapsulate tidy data prinicples because
every row corresponds to a sequence and every column corresponds to a variable.
Hence, we have designed the `plyranges` DSL to have the grammar
and principles of `dplyr`: fluency, consistency, cohesion and endomorphism.
